---
title: "2013 College Scoreboard Analysis"
author: "Zihan Mo"
date: "April 17, 2018"
output:
  html_document:
    fig_height: 4
    fig_width: 6
  word_document: default
---
```{r echo=FALSE}
data = readRDS('college_scorecard_2013.rds')
```
Q1.

a.There are 3312 observations are recorded.

b.There are 2431 colleges are recorded based on main campus counts.

Q2
```{r echo=FALSE, warning=FALSE}
library(knitr)
kable(table(sapply(data, class)))
```

a.There are 51 features in the dataset, and 7 of them are categorical, 15 of them are discrete.

b.There are 5 types of features: character, factor, integer, logic, and numeric. And numeric can be considered as continuous.

Q3

a.There are 23197 missing values in the dataset.

b.avg_sat has the most missing value. 

c.Information related to students privacy like score and family income has more NA. On the contrary, information related to school expense or teacher salary has relatively less NA.

Q4

a.There are 2569 public colleges recorded and 716 private colleges recorded

b.

```{r echo=FALSE}
prop<-prop.table(table(data$highest_degree, data$ownership), margin = 2)
mosaicplot(prop,shade = TRUE,color = TRUE,
           main = "Highest Degree Per College Type",
           xlab = "Highest Degree Awarded", ylab = "Ownership", las=2)
```

Public college does not provide certificate degree which only privat college provides; Graduate degree has more proportion than other degrees.

Q5

a.The average of undergraduate population is 3599.502 and the median is 1295. 

b.Here is the deciles for undergraduate population
```{r echo = FALSE}
kable(quantile(data$undergrad_pop, na.rm = TRUE, probs = seq(0,1,0.1)))
```

c.The undergraduate population has a right skewed distribution.
```{r echo = FALSE}
UG_mean<-mean(data$undergrad_pop, na.rm = TRUE)
UG_med<-median(data$undergrad_pop, na.rm=TRUE)
plot(density(data$undergrad_pop, na.rm = TRUE),
     main="undergraduate population distribution",
     xlim=c(0,170000))
abline(v = UG_mean, col = "red", lwd = 1)
abline(v = UG_med, col = "green", lwd = 1)
legend("topright",legend=c('mean','median'),
       col=c("red","green"),pch="-")
```

d.There is an unusual large undergraduate population which is 166816 in the data. Other campuses only have population below 70000.

Q6
```{r echo = FALSE}
populous_states<-data[data$state%in%c("CA","TX","NY","IL","FL"),]
fstate<-factor(populous_states$state)
boxplot(tuition~fstate, data = populous_states,na.rm=TRUE,
        main = "Tuition distribution w.r.t states",
        xlab = "States", ylab = "Tuition")
```

Based on the boxplots: a. All the states have similar median tuition; b. Gernerally, CA, NY, and IL have higher tuition.

Q7

a.
```{r}
college_sat<-data[,c("name","avg_sat")]
highest_sat<-college_sat$avg_sat
college_sat[which.max(highest_sat),]
```
b.
```{r}
college_UG_pop <- data[, c("name","undergrad_pop","open_admissions")]
largest_UGpop<-college_UG_pop$undergrad_pop
college_UG_pop[which.max(largest_UGpop),]
```
c.
```{r}
fam_income<-data[,c("zip","avg_family_inc","ownership")]
P_fam_income<-subset(fam_income, ownership == "Public")
lowest_fminc<-which.min(P_fam_income$avg_family_inc)
P_fam_income[lowest_fminc,]
```
d.
```{r}
which.max(largest_UGpop)==which.max(data$grad_pop)
```
The campus with the largest undergrad population doesn't have the largest graduate population.

Q8

a.
```{r echo = FALSE}
data2<-subset(data, ownership=="Public"&primary_degree=="Bachelor")
model1<-lm(data2$spend_per_student~data2$revenue_per_student)
plot(data2$spend_per_student, data2$revenue_per_student,
     xlab = "spend_per_student", ylab = "revenue_per_student",
     main = "revenue against spend",las = 1)
abline(model1, col="red", lwd = 2)
```

Revenue_per_student and spend_per_student have positive relationship. There are some outliers in the plot and they may affect the linear regression line.

b.
```{r echo = FALSE, warning=FALSE}
library(plyr)
per_net_income<-data2$revenue_per_student-data2$spend_per_student
total_student_pop<-data2$undergrad_pop+data2$grad_pop
data2$total_net_income<-per_net_income*total_student_pop
data22<-data2[,c("name","total_net_income")]
kable(head(arrange(data22,desc(total_net_income)), n = 5))
```

Q9

Part a

```{r echo=FALSE}
data_group1<-subset(data,data$admission<0.6&data$avg_sat>1200)
data_group2<-subset(data, data$admission>=0.6&data$avg_sat<=1200)
par(mfrow = c(1,2))
plot(data_group1$avg_sat, data_group1$admission,
     main = "group 1", xlab = "avg_sat",
     ylab = "admission rate")
plot(data_group2$avg_sat, data_group2$admission,
     main = "group 2", xlab = "avg_sat",
     ylab = "admission rate")
```

Split the data into two groups: 

1.group 1 has avg_sat > 1200 and the admission rate < 0.6.

2.group 2 has avg_sat <= 1200 and the admission rate >= 0.6.

From the plots above, it's obvious that as avg_sat goes up, admission rate goes down in group 1; However, in group 2, there is no obvious patter. Furthermore, the correlation coefficient of avg_sat and admission rate in group 2 is -0.81, while the the correlation coefficient of avg_sat and admission rate in group 2 is -0.15. The correlation coefficients also clarify the relationships.

PART b

a)

```{r echo=FALSE}
par(mfrow = c(1,2))
boxplot(data_group1$med_10yr_salary, na.rm = TRUE,
             xlab="med_10yr_salary", main = "group1")
boxplot(data_group2$med_10yr_salary,na.rm = TRUE,
     xlab="med_10yr_salary", main = "group2")
```

Based on the density plots above, 75% of students in group 1 earn med_10yr_salary  less than 70000, while 75% of students in group 2 earn med_10yr_salary less than $50000. The med_10yr_salary increases from group 2 to group 1.

b)

```{r echo = FALSE}
data_group1$percen_white_asian1 <- data_group1$race_white+data_group1$race_asian
data_group2$percen_white_asian2 <- data_group2$race_white+data_group2$race_asian
par(mfrow = c(1,2))
boxplot(data_group1$percen_white_asian1, na.rm = TRUE,
     xlab="percent_white_asian", main = "group1")
boxplot(data_group2$percen_white_asian2,na.rm = TRUE,
     xlab="percent_white_asian", main = "group2")
```

In group 1, 75% of the colleges have less than 75% of white and Asian students. While in group two, 75% of the colleges have less than 86% of white and Asian students. The race in group 1 is more diverse than in group 2. The white_Asian percentage increases from group 2 to group 1.

c)
```{r echo = FALSE}
data_group1$percent_grad1<-data_group1$grad_pop/(data_group1$grad_pop+
  data_group1$undergrad_pop)
data_group2$percent_grad2<-data_group2$grad_pop/(data_group2$grad_pop+
  data_group2$undergrad_pop)
par(mfrow = c(1,2))
boxplot(data_group1$percent_grad1, na.rm = TRUE,
        xlab="percent_grad_enrolled", main = "group1")
boxplot(data_group2$percent_grad2, na.rm = TRUE,
        xlab="percent_grad_enrolled", main = "group2")
```

Based on the plots above, colleges in group 1 have higher percentage of graduate students enrolled. percentage of graduate students enrolled increases from group2 to group1. 

PART C

a)
```{r}
sum(data_group1$open_admissions)
sum(data_group2$open_admissions)
```

In both groups, all the colleges are not open admission. Therefore, it's fair to conclude open_admissions is independent of group.

b)
```{r}
num_main_campus1<-sum(data_group1$main_campus)
num_main_campus2<-sum(data_group2$main_campus)
num_main_campus2/830
num_main_campus1/136
```

In group1, 100% of colleges are main campus; in group 2, 96% of colleges are main campus. It's fair to conclude main_campus is independent of group.

c)
```{r echo=FALSE}
tbl1<-table(data_group1$ownership)
tbl2<-table(data_group2$ownership)
percent1<-tbl1/136
percent2<-tbl2/830
par(mfrow=c(1,2))
barplot(percent1)
barplot(percent2)
```

Based on the bar plots above, the percentage of public school in group 2 is twice as much as in group 1. For nonprofit decreases 19% from group 1 to group 2. But for profit almost doesn't change. Therefore, it's fair to conclude ownership is dependent of group. 

d)
```{r echo = FALSE}
tbl3<-table(data_group1$branches)
tbl4<-table(data_group2$branches)
percent3<-tbl3/136
percent4<-tbl4/830
par(mfrow = c(1:2))
barplot(percent3, main = "group1", xlab = "number of branch",
        ylab = "percentage")
barplot(percent4, main = "group2", xlab = "number of branch",
        ylab="percentage")
```

Based on the bar plots above, group 1 and group 2 have similar percentage distribution that 90% of the colleges have one branch. Therefore, it's fair to conclude the variable "branch" is independent of group.

Q10

Part a

```{r echo = FALSE}
par(mfrow=c(1,2))
plot(data$avg_family_inc,data$avg_10yr_salary,
     main = "fam_inc VS 10yr_salary",
     xlab="avg_10yr_salary",ylab="avg_family_inc")
fit1<-lm(data$avg_10yr_salary~data$avg_family_inc)
abline(fit1, col = "green", lwd=2)

data10 <- subset(data, avg_family_inc<=75000)
plot(data10$avg_family_inc,data10$avg_10yr_salary,
     main = "2.fam_inc VS 10yr_salary",
     xlab="avg_10yr_salary",ylab="avg_family_inc")
fit2<-lm(data10$avg_10yr_salary~data10$avg_family_inc)
abline(fit2, col = "red", lwd = 2)
```

For avg_family_inc<=75000(plot on the left), the regression line becomes flatter. The slope includes avg_family_inc > 75000 is 0.247, while the slope not includes avg_family_inc > 75000 is 0.159.

Part b

By adding categorial primary_degree, the adjusted R^2 increseases to 0.2525 from 0.1405.
The levels of the variable is "other", "Certificate", "Associate", "Bachelor", and "Graduate"

###reference
1.Piazza
2.stackoverflow:https://stackoverflow.com/questions/12187891/how-to-get-top-n-companies-from-a-data-frame-in-decreasing-order

###Appendix
```{r, echo=TRUE,eval=FALSE}
data = readRDS('college_scorecard_2013.rds')
1
dim(data)
num_of_obs <- nrow(data)
head(data, n = 4L)
num_of_college <- sum(data$main_campus)

2
str(data)
table(sapply(data, class))
sapply(data, typeof)

3

total_num_NA<-sum(is.na(data))
NA_mat<-is.na(data)
non_NA_mat<-!is.na(data)
NA_counts<-colSums(is.na(data))
sort(NA_counts, decreasing = TRUE)
table(is.na(data$med_family_inc), data$ownership)
table(is.na(data$avg_sat),data$open_admissions)

4
table(factor(data$ownership=="Public",
             labels = c("Public", "Private")))
data$IsPublic<-factor(data$ownership=="Public",
                              labels=c("Private","Public"))
prop<-prop.table(table(data$highest_degree, data$ownership), margin = 2)
mosaicplot(prop,shade = TRUE,color = FALSE,
           main = "Highest Degree Per College Type",
           xlab = "Highest Degree Awarded", ylab = "Ownership", las=2)

5
UG_mean<-mean(data$undergrad_pop, na.rm = TRUE)
UG_med<-median(data$undergrad_pop, na.rm=TRUE)
quantile(data$undergrad_pop, na.rm = TRUE, probs = seq(0,1,0.1))
plot(density(data$undergrad_pop, na.rm = TRUE),
     main="undergrad population distribution",
     xlim=c(0,170000))
abline(v = UG_mean, col = "red", lwd = 1)
abline(v = UG_med, col = "green", lwd = 1)
legend("topright",legend=c('mean','median'),
       col=c("red","green"),pch="-")
boxplot(data$undergrad_pop)
summary(data$undergrad_pop)

6
populous_states<-data[data$state%in%c("CA","TX","NY","IL","FL"),]
fstate<-factor(populous_states$state)
boxplot(tuition~fstate, data = populous_states,na.rm=TRUE,
        main = "Tuition distribution w.r.t states",
        xlab = "States", ylab = "Tuition")

7
a.
college_sat<-data[,c("name","avg_sat")]
highest_sat<-college_sat$avg_sat
college_sat[which.max(highest_sat),]

b
college_UG_pop <- data[, c("name","undergrad_pop","open_admissions")]
largest_UGpop<-college_UG_pop$undergrad_pop
college_UG_pop[which.max(largest_UGpop),]

c
fam_income<-data[,c("zip","avg_family_inc","ownership")]
P_fam_income<-subset(fam_income, ownership == "Public")
lowest_fminc<-which.min(P_fam_income$avg_family_inc)
P_fam_income[lowest_fminc,]

d
which.max(largest_UGpop)==which.max(data$grad_pop)
data[which.max(data$grad_pop),"undergrad_pop"]

8
a
data2<-subset(data, ownership=="Public"&primary_degree=="Bachelor")
model1<-lm(data2$spend_per_student~data2$revenue_per_student)
plot(data2$spend_per_student, data2$revenue_per_student,
     xlab = "spend_per_student", ylab = "revenue_per_student",
     main = "revenue against spend",las = 1)
abline(model1, col="red", lwd = 2)

b
per_net_income<-data2$revenue_per_student-data2$spend_per_student
total_student_pop<-data2$undergrad_pop+data2$grad_pop
data2$total_net_income<-per_net_income*total_student_pop
data22<-data2[,c("name","total_net_income")]
head(arrange(data22,desc(total_net_income)), n = 5)

9
part A
data_group1<-subset(data, data$admission<0.6&data$avg_sat>1200)
data_group2<-subset(data, data$admission>=0.6&data$avg_sat<=1200)
par(mfrow = c(1,2))
plot(data_group1$avg_sat, data_group1$admission,
     main = "group 1", xlab = "avg_sat",
     ylab = "admission rate")
plot(data_group2$avg_sat, data_group2$admission,
     main = "group 2", xlab = "avg_sat",
     ylab = "admission rate")
r_group2<-cor(data_group2$avg_sat, data_group2$admission)
r_group1<-cor(data_group1$avg_sat, data_group1$admission)

PART B
a)
par(mfrow = c(1,2))
boxplot(data_group1$med_10yr_salary, na.rm = TRUE,
             xlab="med_10yr_salary", main = "group1")
boxplot(data_group2$med_10yr_salary,na.rm = TRUE,
     xlab="med_10yr_salary", main = "group2")

b)
data_group1$percen_white_asian1 <- data_group1$race_white+data_group1$race_asian
data_group2$percen_white_asian2 <- data_group2$race_white+data_group2$race_asian
par(mfrow = c(1,2))
boxplot(data_group1$percen_white_asian1, na.rm = TRUE,
     xlab="percen_white_asian", main = "group1")
boxplot(data_group2$percen_white_asian2,na.rm = TRUE,
     xlab="percen_white_asian", main = "group2")

c)
data_group1$percent_grad1<-data_group1$grad_pop/(data_group1$grad_pop+
  data_group1$undergrad_pop)
data_group2$percent_grad2<-data_group2$grad_pop/(data_group2$grad_pop+
  data_group2$undergrad_pop)
par(mfrow = c(1,2))
boxplot(data_group1$percent_grad1, na.rm = TRUE,
        xlab="percent_grad_enrolled", main = "group1")
boxplot(data_group2$percent_grad2, na.rm = TRUE,
        xlab="percent_grad_enrolled", main = "group2")

PART C
a)
num_open_ad1<-sum(data_group1$open_admissions)
num_open_ad2<-sum(data_group2$open_admissions)

b)
num_main_campus1<-sum(data_group1$main_campus)
num_main_campus2<-sum(data_group2$main_campus)
percentage2<-num_main_campus2/830
percentage1<-num_main_campus1/136

c)
tbl1<-table(data_group1$ownership)
tbl2<-table(data_group2$ownership)
percent1<-tbl1/136
percent2<-tbl2/830
par(mfrow=c(1,2))
barplot(percent1, main = "group1", xlab = "ownership",
        ylab = "percentage")
barplot(percent2, main = "group2", xlab = "ownership",
        ylab="percentage")

d)
tbl3<-table(data_group1$branches)
tbl4<-table(data_group2$branches)
percent3<-tbl3/136
percent4<-tbl4/830
barplot(percent3, main = "group1", xlab = "number of branch",
        ylab = "percentage")
barplot(percent4, main = "group2", xlab = "number of branch",
        ylab="percentage")

10
part A
par(mfrow=c(1,2))
plot(data$avg_family_inc,data$avg_10yr_salary,
     main = "fam_inc VS 10yr_salary",
     xlab="avg_10yr_salary",ylab="avg_family_inc")
fit1<-lm(data$avg_10yr_salary~data$avg_family_inc)
abline(fit1, col = "green", lwd=2)
data10 <- subset(data, avg_family_inc<=75000)
plot(data10$avg_family_inc,data10$avg_10yr_salary,
     main = "fam_inc VS 10yr_salary",
     xlab="avg_10yr_salary",ylab="avg_family_inc")
fit2<-lm(data10$avg_10yr_salary~data10$avg_family_inc)
abline(fit2, col = "red", lwd = 2)
summary(fit1)
summary(fit2)

PART B
fit3<-lm(data$avg_10yr_salary~data$avg_family_inc+data$primary_degree)
abline(fit3, col = "blue", lwd=2)
levels(data$primary_degree)
```