---
title: "STA 141A HW2"
author: "Zihan Mo 914998952"
date: "April 29, 2018"
output:
  html_document: default
  word_document: default
---
```{r echo=FALSE}
data<-readRDS('housing.rds')
```
```{r echo=FALSE, message=FALSE}
library(dplyr)
library(lubridate)
library(ggplot2)
library(MASS)
library(maps)
library(treemap)
```
Q1 Justified by code in the appendix

Q2
The timespan of the housing sales cover from 2003-04-27 to 2006-06-04;
The timespan of the construction year of houses cover from 1885 to 2005.

Q3
```{r echo = FALSE}
data<-mutate(data,yr=year(data$date),mon=month(data$date))
ggplot(data=data, aes(y=price,x=factor(mon),group=yr,color=factor(yr)))+
  stat_summary(fun.y = "mean", geom = "line")+
  xlab("Month")+ylab("Avg Housing Price")+labs(color="year")+
  ggtitle("Avg Price Over Time")+theme_bw()
ggplot(data=data, aes(mon))+
  scale_x_continuous(breaks=seq(1,12,1))+
  geom_histogram(aes(fill=factor(yr)),binwidth = 0.5)+
  stat_bin(binwidth = 0.5, geom="text", colour="black", size = 3.5, 
           aes(label=..count.., group = yr), vjust = 1)+
  ylab("Sales")+ggtitle("Sales Over Time")+
  scale_fill_discrete(name="Year")
```

Q4
```{r echo=FALSE}
data$county<-sub("county","County",data$county)
data$county<-sub("San Franciscoe","San Francisco",data$county)
data$county<-sub("Alpine","San Francisco",data$county)
data$county<-as.factor(data$county)
dat4<-filter(data, year(date)%in%c("2005","2004","2003"))
dat4$br<-cut(as.numeric(dat4$br), breaks = c(0,1,2,3,28),
             labels=c("1","2","3","4+"))
dat4$yr<-as.factor(year(dat4$date))
ggplot(data = na.omit(dat4), aes(y=price,x=yr,group=br,color = br))+
  stat_summary(fun.y = "mean", geom = "line")+
  facet_wrap(~county)+
  theme_bw()+xlab("year")+ylab("avg_price")+
  ggtitle("Avg_price w.r.t Year/Br/County")+theme_classic()
```

Q5

Only Vallejo city has sales in more than one county, which are Napa County and Solano County.

Q6

After taking out extreme outliers, like price equals zero and the building size larger than 10000, we have plots. Based on the plots below, the linear regression model seems provide a good fit to the data. But the QQ-plot and the residuals plot imply the nomality assumption isn't met and the variance of residual is not constant.
```{r echo=FALSE}
dat6<-subset(data,price>0&price<6000000&bsqft<10000)
par(mfrow=c(1,3))
plot(dat6$bsqft, dat6$price,xlab = "bsqft",ylab="price",
     main="scatter plot")
model1<-lm(dat6$price~dat6$bsqft)
abline(model1,col="red",lwd=2)
plot(model1,which=2)
plot(model1,which=1)
```

Based on the QQ-plot below, after BoxCox transformation, the QQ-plot roughly follows a straight line.
```{r echo=FALSE}
model2<-lm(dat6$price^(-0.061)~dat6$bsqft)
qqnorm(model2$residuals)
```

Q7
```{r}
model3<-lm(data$price~data$bsqft+data$lsqft)
model3_est<-coef(summary(model3))[,1]
model3_se<-coef(summary(model3))[,2]
t_star<-(model3_est[2]-model3_est[3])/(model3_se[2])
critical_val<-qt(0.99,19997)
t_star>=critical_val
```
H Null:; H1:;
Using t test, because t_star is greater than critical_val, conclude null hypothesis.

Q8
```{r echo=FALSE}
plot(1, type="n", xlab="bsqft", ylab="price", 
     xlim=range(data$bsqft,na.rm = TRUE),
     ylim=range(data$price,na.rm = TRUE),
     main="regression models")
split_county<-split(data,data$county)
models<-lapply(split_county,function(x)lm(price~bsqft,x))
regression_lines<-sapply(1:9, function(i) abline(coef(models[[i]]), col=i))
legend("bottomright",names(table(data$county)),
       col=1:9,lty=rep(1,9),cex = 0.7)
```

It's obviouse that the simple regression lines with respect to different county are not parallel. Therefore, we can conclude the regression lines depend on county.

Q9
```{r echo=F}
sales<-aggregate(price~city+county,data,length)
avg_price<-aggregate(price~city+county,data,mean)
colnames(sales)[3]<-"sales"
colnames(avg_price)[3]<-"avg_price"
price_sales<-merge(sales,avg_price,by=c("city","county"))
price_sales<-price_sales%>%
  group_by(county)%>%
  top_n(n=3L,wt=sales)
treemap(price_sales,
        index = c("county","city"),
        vColor="sales",
        vSize ="avg_price" ,
        type="value",
        title='top3 sales with avg_price in differetn county',
        align.labels=list(c("left","top"),
                          c("center","center")),
        format.legend = list(scientific = FALSE, big.mark = " "))
```

Q10
```{r echo=F}
SFdata<-subset(data,data$city=="San Francisco")
SFdata$long2<-round(SFdata$long,2)
SFdata$lat2<-round(SFdata$lat,2)
#factorize the long,lat ranges,form cells.
long2_range<-range(SFdata$long2,na.rm=TRUE)
long_seq<-seq(long2_range[1],long2_range[2],.01)
SFdata$longF<-factor(SFdata$long2,levels=long_seq)

lat2_range<-range(SFdata$lat2,na.rm = T)
lat_seq<-seq(lat2_range[1],lat2_range[2],.01)
SFdata$latF<-factor(SFdata$lat2,levels=lat_seq)
SF_agg<-aggregate(price~latF+longF,SFdata,
                  function(x)c(mean(x),length(x)),drop=FALSE)

SFhouse_prices<-matrix(SF_agg$price[,1],
                     nlevels(SFdata$longF),
                     nlevels(SFdata$latF),byrow=TRUE)
SFhouse_prices[SFhouse_prices==0]<-NA
SFhouse_sales<-matrix(SF_agg$price[,2],
                     nlevels(SFdata$longF),
                     nlevels(SFdata$latF),byrow=TRUE)
SFhouse_sales[SFhouse_sales==0]<-NA

sf_border=map('county','california,san francisco',plot = F)
image(x=long_seq,y=lat_seq,z=SFhouse_prices,xlim=c(-122.56,-122.3),
      ylim = c(37.7,37.82),
      xlab = "Longitude", ylab = "Latitude", 
      main = "SF Heatmap of Average Housing Prices")
lines(sf_border$x,sf_border$y)


image(x=long_seq,y=lat_seq,z=SFhouse_sales,xlim=c(-122.56,-122.3),
      ylim = c(37.7,37.82),
      xlab = "Longitude", ylab = "Latitude", 
      main = "SF Heatmap of Housing Sales")
lines(sf_border$x,sf_border$y)
```





